Gompertz growth as infinitely correlated non-local growth

### 2023-10-09 arxiv2/new_draft.tex
update

\usepackage{arxiv}
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{lipsum}   % Can be removed after putting your text content
\usepackage{graphicx}
\usepackage[square,compress,numbers]{natbib}
\usepackage{doi}
\usepackage{amsmath}
\usepackage{float}
\usepackage{lineno}
% \linenumbers

@@

\author{Matz A. Haugen, \\
Independent Reseacher, Oslo\\
\texttt{matzhaugen@gmail.com} \\
  \And
  Dorothea Gilbert \\
  Independent Reseacher, Oslo\\
}

@@

\begin{abstract}
Logistic and Gompertz growth has traditionally been connected by augmenting the logistic model with an extra parameter to obtain what is known as $\theta$-logistic growth or the Richards model for growth. In spite of this bridge, the biological foundation of Gompertz growth is only vaguely understood. 
To add to our biological understanding, we show that there is another way to connect the two growth modes by adding higher order terms in the growth equation, and present both a macroscopic and microscopic representation of growth where the latter involves a network of entities. 
The microscopic representation elucidates the source of Gompertz growth as the potential abundance and not the realized abundance, and an example of such a network is presented. 
For finite higher order terms we show that the growth equation can be represented by Gauss' Hypergeometric function, and only for infinitely many terms does the abundance lend itself conveniently to a log-transformation wherein the log-abundance growth rate decreases linearly, the trademark of Gompertz growth. We thus equate Gompertz growth with correlated microscopic entities that exhibit long-range or non-local interaction. 
This conclusion adds to the interpretation of biological systems that undergo Gompertz growth where long-range interaction could be interpreted as the source of growth being external to the individual entities, like a field that stimulates growth simultaneously across all entities. 
This field could be interpreted as an environmental stressor like air-pollution, age, radiation, etc, or an environmental stimulant like an electromagnetic field. 
Our new discovery allows for a unification of well-known growth models as follows: time independent and uncorrelated growth yields the exponential growth function, time dependent and uncorrelated growth yields Gaussian growth, time independent and pairwise correlated growth yields the logistic function and time independent and infinitely correlated growth yields the Gompertz function.

@@

% keywords can be removed
\keywords{Gompertz \and Coherence \and Network Analysis \and Stochastic}

@@

\begin{equation}
\frac{d \ln{x}}{dt} = \sum_{i=1}^K s^i/i.
\end{equation}
Using standard integrals we see that this sum amounts to a Gompertz growth term adjusted by a Hypergeometric function,
\begin{eqnarray*}
  \sum_{i=1}^K s^i/i &=& \int_s \sum_{i=0}^{K-1} s^i ds\\
  &=&\int_s \frac{1-s^K}{1-s} ds\\
  &=&- \frac{{}_{1}F_{2}({1,K+1,K+2;s})}{K+1}s^{K+1} - \ln(1-s),
\end{eqnarray*}
where ${}_1F_{2}$ is Gauss' Hypergeometric function, and integration limits are indefinite. In summary, the modified logistic function emerges as
 \begin{equation}
 \label{eq:modLogistic2}
\frac{d}{dt}\ln{x(t)} = - \frac{{}_{1}F_{2}({1,K+1,K+2;1-x(t)})}{K+1}(1-x(t))^{K+1} - \ln{x(t)},
 \end{equation}
 which represents the novel bridge between logistic and Gompertzian growth. Note that for $K=1$, this relative growth rate becomes the standard logistic function, while increasing the number of higher order terms will move the equation closer to the Gompertz curve as the Hypergeometric term approaches zero. This modification to the logistic function does not have a 1-to-1 correspondence with the Richards model, but serves instead as an alternative when encoding non-linear or collaborative growth effects (see Fig. \ref{fig:hypergeometric_plain}).

>>>>

\usepackage{arxiv}
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{lipsum}   % Can be removed after putting your text content
\usepackage{graphicx}
\usepackage[square,compress,numbers]{natbib}
\usepackage{doi}
\usepackage{amsmath}
\usepackage{float}
\usepackage{lineno}
\linenumbers

@@

\author{Matz A. Haugen\footnote{hi}, \\
Independent Reseacher, Oslo\\
\texttt{matzhaugen@gmail.com} \\
  \And
  Dorothea Gilbert \\
  Independent Reseacher, Oslo\\
}

@@

\begin{abstract}
Logistic and Gompertz growth has traditionally been connected by augmenting the logistic model with an extra parameter to obtain what is known as $\theta$-logistic growth or the Richards model for growth. In spite of this bridge, the biological foundation of Gompertz growth is only vaguely understood. 
To add to this biological foundation, we show that there is another way to connect the two growth modes by adding higher order terms in the growth equation, and present both a macroscopic and microscopic representation of growth where the latter involves a network of entities. 
The microscopic representation elucidates the source of Gompertz growth as the potential abundance and not the realized abundance, and an example of such a network is presented. 
For finite higher order terms we show that the growth equation can be represented by Gauss' Hypergeometric function, and only for infinitely many terms does the abundance lend itself conveniently to a log-transformation wherein the log-abundance growth rate decreases linearly, the trademark of Gompertz growth. We thus equate Gompertz growth with correlated microscopic entities that exhibit long-range or non-local interaction. 
This conclusion adds to the interpretation of biological systems that undergo Gompertz growth where long-range interaction could be interpreted as the source of growth being external to the individual entities, like a field that stimulates growth simultaneously across all entities. 
This field could be interpreted as an environmental stressor like air-pollution, age, radiation, etc, or an environmental stimulant like an electromagnetic field. 
Our new discovery allows for a unification of well-known growth models as follows: time independent and uncorrelated growth yields the exponential growth function, time dependent and uncorrelated growth yields Gaussian growth, time independent and pairwise correlated growth yields the logistic function and time independent and infinitely correlated growth yields the Gompertz function.

@@

% keywords can be removed
\keywords{Gompertz \and Coherence \and Network Analysis \and Self-organization}

\section*{Highlights}
\begin{itemize}
\item Logistic and Gompertz growth is connected with a novel framework based on microscopic principles.
\item Gompertz growth is connected with ubiquitous growth-activating field justified by physical and mathematical principles.
\item Growth processes is put in a larger contextual framework based on varying dependence in spacetime, including the Exponential, Gaussian and Poisson distributions. 
\end{itemize}

@@

\begin{equation}
\frac{d \ln{x}}{dt} = \sum_{i=1}^K s^i/i.
\end{equation}
Using standard integrals we see that this sum amounts to a Gompertz growth term adjusted by a Hypergeometric function,
\begin{eqnarray*}
  \sum_{i=1}^K s^i/i &=& \int_s \sum_{i=0}^{K-1} s^i ds\\
  &=&\int_s \frac{1-s^K}{1-s} ds\\
  &=&- \frac{{}_{1}F_{2}({1,K+1,K+2;s})}{K+1}s^{K+1} - \ln(1-s),
\end{eqnarray*}
where ${}_1F_{2}$ is Gauss' Hypergeometric function, and integration limits are indefinite. In summary, the modified logistic function emerges as
 \begin{equation}
 \label{eq:modLogistic2}
\frac{d}{dt}\ln{x(t)} = - \frac{{}_{1}F_{2}({1,K+1,K+2;1-x(t)})}{K+1}(1-x(t))^{K+1} - \ln{x(t)},
 \end{equation}
 which represents the novel bridge between logistic and Gompertz growth. Note that for $K=1$, this relative growth rate becomes the standard logistic function, while increasing the number of higher order terms will move the equation closer to the Gompertz curve as the Hypergeometric term approaches zero. This modification to the logistic function does not have a 1-to-1 correspondence with the Richards model, but serves instead as an alternative when encoding non-linear or collaborative growth effects (see Fig. \ref{fig:hypergeometric_plain}).
